Combinatorial Proofs and Decomposition Theorems for First-order Logic\vspace{-3ex}

### 2021-01-25 main.tex
LK

%All statements in this section are proved by rule permutation arguments.
\begin{thm}\label{thm:LK1-decompose}
  Let $\Gamma$ be a sequent. If\/ $\sqn\Gamma$ is provable in $\FOLK$ (as depicted on the left below) then there is a sequent
  $\Gamma'$ such that there is a derivation as shown on the right below:\vadjust{\vskip-2ex}
  \begin{equation*}
    \vlderivation{
        \vltrl{}{\FOLK}{\Deri}{\sqns{\Gamma}}{
          \vlhy{\quad}}{
          \vlhy{}}{
          \vlhy{}}}
    \qquad
    \Longrightarrow
    \qquad
    \vlderivation{
      \vlde{\set{\wrD,\cD,\fequ}}{\Deri_2}{\sqns{\form{\Gamma}}}{
        \vltrl{}{\FOMLL}{\Deri_1}{\sqns{\Gamma'}}{
          \vlhy{\quad}}{
          \vlhy{}}{
          \vlhy{}}}}
  \end{equation*}
\end{thm}

\begin{proof}

By transforming each sequent of $\Deri'_2$ into its corresponding formula, and by
considering the following rewriting rule:
\begin{equation*}
\vlderivation{
  \vlin{\weakr}{}{\sqns{\Gamma, A}}{
    \vlhy{\sqns{\Gamma}}}}
\leadsto
\vlderivation{
  \vlin{\wrD}{}{\sqns{\form{\Gamma} \vlor A}}{
    \vlhy{\sqns{\form{\Gamma}}}}}
\end{equation*},
we obtain a derivation
\begin{equation*}
\vlderivation{
  \vlde{\set{\wrD,\cD,\fequ}}{\Deri_2}{\sqns{\form{\Gamma}}}{
    \vltrl{}{\FOMLL}{\Deri_1}{\sqns{\Gamma'}}{
      \vlhy{\quad}}{
      \vlhy{}}{
      \vlhy{}}}}
\end{equation*}
where $\Gamma' = \form{\Gamma_0}$ and $\Deri_1$ can be obtained from $\Deri'_1$
by applying the $\vlor$ rule.
\tocheck{}
\juihsuan{This might be a bit long...}
\end{proof}

@@

>>>>

%All statements in this section are proved by rule permutation arguments.
\begin{thm}\label{thm:LK1-decompose}
  Let $\Gamma$ be a sequent. If\/ $\sqn\Gamma$ is provable in $\FOLK$ (as depicted on the left below) then there is a sequent
  $\Gamma'$ not containing any $\fff$, such that there is a derivation as shown on the right below:\vadjust{\vskip-2ex}
  \begin{equation*}
    \vlderivation{
        \vltrl{}{\FOLK}{\Deri}{\sqns{\Gamma}}{
          \vlhy{\quad}}{
          \vlhy{}}{
          \vlhy{}}}
    \qquad
    \Longrightarrow
    \qquad
    \vlderivation{
      \vlde{\set{\wrD,\cD,\fequ}}{\Deri_2}{\sqns{\form{\Gamma}}}{
        \vltrl{}{\FOMLL}{\Deri_1}{\sqns{\form{\Gamma'}}}{
          \vlhy{\quad}}{
          \vlhy{}}{
          \vlhy{}}}}
  \end{equation*}
\end{thm}

\begin{proof}
  First, we can replace every instance of the $\fff$-rule by
  $\weakr$. Then the instances of $\weakr$ and $\conr$ are replaced by
  $\wD$ and $\cD$, which can then be permuted down. (Details are in
  Appendix~\ref{app:LK1-decompose}.
\end{proof}

@@

\subsection{Proof of Theorem~\ref{thm:LK1-decompose}}
\label{app:LK1-decompose}

\begin{proof}[Proof of Theorem~\ref{thm:LK1-decompose}]

By transforming each sequent of $\Deri'_2$ into its corresponding formula, and by
considering the following rewriting rule:
\begin{equation*}
\vlderivation{
  \vlin{\weakr}{}{\sqns{\Gamma, A}}{
    \vlhy{\sqns{\Gamma}}}}
\leadsto
\vlderivation{
  \vlin{\wrD}{}{\sqns{\form{\Gamma} \vlor A}}{
    \vlhy{\sqns{\form{\Gamma}}}}}
\end{equation*},
we obtain a derivation
\begin{equation*}
\vlderivation{
  \vlde{\set{\wrD,\cD,\fequ}}{\Deri_2}{\sqns{\form{\Gamma}}}{
    \vltrl{}{\FOMLL}{\Deri_1}{\sqns{\Gamma'}}{
      \vlhy{\quad}}{
      \vlhy{}}{
      \vlhy{}}}}
\end{equation*}
where $\Gamma' = \form{\Gamma_0}$ and $\Deri_1$ can be obtained from $\Deri'_1$
by applying the $\vlor$ rule.
\end{proof}
